---
title: "NYC_Crime_Data_V2"
author: "J. Luciano"
date: '2022-04-17'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYC Crime Data
The report contains data from the following website: https://data.cityofnewyork.us.

Overall goal was to identify if there is a high correlation between the amount of arrests and shootings.

I have downloaded a CSV file of Arrest Data and Shooting Data. 

I have analyzed the amount of shooting by Borough (Boro) and Precinct.

I have also included a linear model with shooting being the target value and it being based on the the number of arrests.

Note - I have summed all historic data.




Libraries uses
```{r libraries_used}
library(RCurl)
library(tidyverse)
library(dplyr)

```

Reading in the data from 2 csv files NYC Shooting Data and NYC Arrest Data
```{r get_nyc_crime_data, message = FALSE}
NYC_Shooting_data <- read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")
NYC_Arrest_Data <- read.csv("https://data.cityofnewyork.us/api/views/8h9b-rp9u/rows.csv?accessType=DOWNLOAD", header=TRUE)

```

Creating a Bar Column with total shooting with color views of each borough (BORO)
```{r Creating_Bar_Plot_NYC_Shooting_By_Boro }
NYC_BP <- ggplot(NYC_Shooting_data, aes(x="", y = INCIDENT_KEY ,fill=BORO))+
geom_bar(width = 1, stat = "identity")
NYC_BP

```

Creating a Pie Chart Column with total shooting with color views of each borough (BORO)
```{r Creating_Pie_Chart_NYC_Shooting_By_Boro}
NYC_Pie <- NYC_BP + coord_polar("y",start=0)

NYC_Pie

```

Creating a count of shooting, group by data is based on Precinct
Data is shown

```{r Count_by_precint_Shooting_data}
NYC_Shooting_BY_PRECINCT <- NYC_Shooting_data %>% group_by(PRECINCT) %>% tally()
NYC_Shooting_BY_PRECINCT  <- NYC_Shooting_BY_PRECINCT  %>%
rename(Number_Of_Shootings  = n)
NYC_Shooting_BY_PRECINCT

```
bREAK 2


count of arrest by precicint
we also changed the name of arrest_precinct columnt to precinct
data is shown
```{r Count_by_precint_Arrest_data }

NYC_Arrest_BY_PRECINCTS <- NYC_Arrest_Data %>% group_by(ARREST_PRECINCT) %>% tally()

NYC_Arrest_BY_PRECINCTS  <- NYC_Arrest_BY_PRECINCTS  %>%
rename(PRECINCT  = ARREST_PRECINCT)

NYC_Arrest_BY_PRECINCTS  <- NYC_Arrest_BY_PRECINCTS  %>%
rename(Number_Of_Arrests  = n)
NYC_Arrest_BY_PRECINCTS
```


Merge NYC_Shooting_BY_PRECINCT and NYC_Arrest_By_Precinct data frames (join them together)
Data is shown
```{r Merge_Data }

NYC_Combined_Crime_Data = merge(x=NYC_Arrest_BY_PRECINCTS ,y=NYC_Shooting_BY_PRECINCT  ,by="PRECINCT",all.y=TRUE)
NYC_Combined_Crime_Data

```

Building Predictive Linear Model with NYC_Combined_Data
Will look to predict the amount shootings based on how many arrests are made
```{r Linear_Model }

mod <- lm(Number_Of_Shootings ~ Number_Of_Arrests , data = NYC_Combined_Crime_Data)
summary(mod)



```

Building LInear Graphs that first plot number of arrests compared to shooting 
Then we add the linear model in with a red line.

```{r Linear_Graph }


shootings.graph<-ggplot(NYC_Combined_Crime_Data, aes(x=Number_Of_Arrests, y=Number_Of_Shootings))+ geom_point()
```

```{r Linear_Graph2 }
shootings.graph <- shootings.graph + geom_smooth(method="lm", col="red")
shootings.graph


```
Closing wording, I did have a bias, as I would have assumed that the there would be a high correlation of shootings where a high number of arrests were made.

Also, my assumption is that in Staten Island which has a more suburban feel, there would be less shootings per arrest.

Based on the data there is a high correlation when comparing arrests to shootings in each precinct.

Ultimately, there is a probably high causation based on weatlth, population density, access to education and other factors that would need to be included in the analysis.

Higher crimes in itself will lead to more arrests and shootings, thus, the correlation makes sense.

sessionInfo()
